---
title: "p8105_hw2_zc2610"
author: "Zuoqiao Cui"
date: "2022-10-01"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1

```{r}
NYC_transit_data = read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line,station_name,station_latitude,station_longitude,starts_with("route"),entry, exit_only, vending, entrance_type, ada) %>% 
  mutate(
    entry = recode(entry,"YES" = TRUE,"NO" = FALSE),
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)
  )
```

```{r}
num_of_row = nrow(NYC_transit_data)
num_of_column = ncol(NYC_transit_data)
```


Description

1. There are these variables: `r colnames(NYC_transit_data)`

2. The process of tidying data is to import, update variable names, make all data type of column route to be the same: character, and selects the columns that will be used in later parts fo this problem. Then, I update `entry` from `yes` / `no` to a logical variable.

3.There are totally `r nrow(NYC_transit_data)` rows and `r ncol(NYC_transit_data)` columns.

4. I think it is not tidy enough since there are several columns regarding route and they can be combined to a column "route" with different numbers as values


Calculate the number of combinations of different station names and lines
```{r}
num_of_combination = NYC_transit_data %>% 
  select(station_name,line) %>% 
  distinct
```
Description
There are totally `r nrow(num_of_combination)` distinct stations

Calculate the number of stations where adas are compliant
```{r}
num_of_combination_ada = NYC_transit_data %>%
filter(ada == TRUE) %>% 
  select(station_name,line) %>% 
  distinct
```
Description
There are totally `r nrow(num_of_combination_ada)` distinct stations where adas are compliant


calculate the proportion of station entrances / exits without vending allow entrance
```{r}
proportion = NYC_transit_data %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
proportion
```
Description

The proportion of station entrances / exits without vending allow entrance is `r proportion`

reform route number and route name
calculate the number of distinct stations that serve the A train
```{r}
num_a = NYC_transit_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct
```
Description

There are totally `r nrow(num_a)` distinct stations serve the A train


calculate the number of stations that serve the A train and are ADA compliant
```{r}
num_a_ada = NYC_transit_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
Description

There are totally `r nrow(num_a_ada)` distinct stations that serve the A train and are ADA compliant

## Problem 2

Import, clean, and organize the data for Mr Trash Wheel

```{r}
MrTrashWheel_df = 
    read_excel("./data/Trash Wheel Collection Data.xlsx",sheet = "Mr. Trash Wheel",range = "A2:N550") %>% 
    janitor::clean_names() %>% 
    drop_na(dumpster) %>% 
    mutate(
      dumpster = as.character(dumpster),
      sports_balls = as.integer(round(sports_balls)),
      year = as.double(year)
    )
MrTrashWheel_df
```
Import, clean, and organize the data for Professor Trash Wheel

```{r}
ProfessorTrashWheel_df = 
    read_excel("./data/Trash Wheel Collection Data.xlsx",sheet = "Professor Trash Wheel",range = "A2:M97") %>% 
    janitor::clean_names() %>% 
    drop_na(dumpster) %>% 
    mutate(
      dumpster = as.character(dumpster)
    )
ProfessorTrashWheel_df
```

Combine two sheets

```{r}
MrTrashWheel_df = MrTrashWheel_df %>% 
 mutate(type = "Mr. Trash Wheel")
ProfessorTrashWheel_df = ProfessorTrashWheel_df %>% 
  mutate(type = "Professor Trash Wheel")
TrashWheel_tidy = bind_rows(MrTrashWheel_df,ProfessorTrashWheel_df) %>% 
  janitor::clean_names() %>% 
  mutate(
    year = as.integer(year)
  )
TrashWheel_tidy
```
Number of observations and variables in Mr Trash Wheel dataset
```{r}
num_of_MrTrashWheel_observations = nrow(MrTrashWheel_df)
num_of_MrTrashWheel_variables = ncol(MrTrashWheel_df)
num_of_MrTrashWheel_observations
num_of_MrTrashWheel_variables
```

Number of observations and variables in Mr Trash Wheel dataset
```{r}
num_of_ProfessorTrashWheel_observations = nrow(ProfessorTrashWheel_df)
num_of_ProfessorTrashWheel_variables = ncol(ProfessorTrashWheel_df)
num_of_ProfessorTrashWheel_observations
num_of_ProfessorTrashWheel_variables
```

Number of observations and variables in Mr Trash_tidy date frame
```{r}
num_of_TrashWheel_tidy_observations = nrow(TrashWheel_tidy) 
num_of_TrashWheel_tidy_variables = ncol(TrashWheel_tidy)
num_of_TrashWheel_tidy_observations
num_of_TrashWheel_tidy_variables
```
Calculate the total weight of trash collected by professor trash wheel
```{r}
professor_total_weight = sum(filter(TrashWheel_tidy,type == "Professor Trash Wheel")$weight_tons)
professor_total_weight
```

Calculate the total number of sports bals collected by Mr trash wheel
```{r}
Mr_total_sportsballs = sum((filter(TrashWheel_tidy,type == "Mr. Trash Wheel",year == 2020)$sports_balls))
Mr_total_sportsballs
```

Description

1.There are totally `r nrow(MrTrashWheel_df)` observations and `r ncol(MrTrashWheel_df)` variables in the Mr Trash Wheel dataset. 

2.There are totally `r nrow(ProfessorTrashWheel_df)` observations and `r ncol(ProfessorTrashWheel_df)` variables in the Professor Trash Wheel dataset. 

3.There are totally `r nrow(TrashWheel_tidy)` observations and `r ncol(TrashWheel_tidy)` variables in the final combined dataset. 

4.Key variables contain "dumpster","year","weight_tons","sports_balls" and so on. All variables are as followed: `r colnames(TrashWheel_tidy)`

5.The total weight of trash collected by Professor Trash Wheel is `r sum(filter(TrashWheel_tidy,type == "Professor Trash Wheel")$weight_tons)`

6.The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(filter(TrashWheel_tidy,type == "Mr. Trash Wheel",year == 2020)$sports_balls)`

## Problem 3

1. Separate date into year, month and day
2. Replace month number with month name
```{r}
pols_month_df = 
    read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
    janitor::clean_names() %>% 
    separate(mon,into = c("year","month","day"),sep = "-") %>% 
  mutate(
    month = recode(month, "01" = "January",
                  "02" = "February",
                  "03" = "March",
                  "04" = "April",
                  "05" = "May",
                  "06" = "June",
                  "07" = "July",
                  "08" = "Augest",
                  "09" = "September",
                  "10" = "October",
                  "11" = "November", 
                  "12" = "December"
    )
  )
pols_month_df
```
create a president column using values prez_gop and prez_dem
```{r}
pols_month_df = pols_month_df %>% 
  mutate(
    president = ifelse(prez_gop == 1, "gop",ifelse(prez_gop == 0,"dem","special case")))
  pols_month_df
```

remove prez_gop,prez_dem and day variables
```{r}
pols_month_df = pols_month_df %>% 
select(-prez_dem,-prez_gop,-day)
pols_month_df
```

separate date into year, month, day
```{r}
snps_df = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
    separate(date,into = c("month","day","year"),sep = "/")
snps_df
```

arrange year,month and organize them to be leading columns
```{r}
snps_df = snps_df %>% 
mutate(
  year = as.integer(year),
  year = ifelse(year < 22,year + 2000,year + 1900)
) 
  snps_df = mutate(snps_df,
    year = as.character(year)
  ) %>% 
  arrange(year,month) %>% 
  relocate(year,month)
  snps_df
```
switch unemployment dataset from “wide” to “long” format
```{r}
unemployment_df = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month", 
    values_to = "percentage of unemployment") %>% 
  mutate(
    year = as.character(year)
  )
unemployment_df
```

merge snps dataset into pols dataset and then merge unemployment dataset into them
```{r}
snps_pols_merge = left_join(pols_month_df,snps_df,by = c("year","month"))
merge_unemployment = left_join(snps_pols_merge,unemployment_df,by = c("year","month"))
snps_pols_merge
merge_unemployment
```
illustration:
since the requirement is merge snps into pols, I think all variables of in pols dataset are more important so that need to be reserved.Thus, I choose to use left_join. To reserve all variables, I think full_join can also be used.

Description

1. pols-month dataset contains variables as followed: `r colnames(pols_month_df)`. There are totally `r nrow(pols_month_df)` observations and `r ncol(pols_month_df)` variables in this dataset. The range of year is from `r min(pull(pols_month_df, year))` to `r max(pull(pols_month_df, year))`. Key variables should be year, month and president.

2. snps dataset contains variables as followed: `r colnames(snps_df)`. There are totally `r nrow(snps_df)` observations and `r ncol(snps_df)` variables in this dataset. The range of year is from `r min(pull(snps_df, year))` to `r max(pull(snps_df, year))`. Key variables should be year and month.

3. unemployment dataset contains variables as followed: `r colnames(unemployment_df)`. There are totally `r nrow(unemployment_df)` observations and `r ncol(unemployment_df)` variables in this dataset. The range of year is from `r min(pull(unemployment_df, year))` to `r max(pull(unemployment_df, year))`. Key variables should be year and month.

4. snps_pols_merge dataset is formed by merging snps into pols. This dataset contains variables as followed: `r colnames(snps_pols_merge)`. There are totally `r nrow(snps_pols_merge)` observations and `r ncol(snps_pols_merge)` variables in this dataset. The range of year is from `r min(pull(snps_pols_merge, year))` to `r max(pull(snps_pols_merge, year))`. Key variables should be year and month since snps and pols are joined by these two variables

5. merge_unemployment dataset is formed by merging unemployment into snps_pols_merge. This dataset contains variables as followed: `r colnames(merge_unemployment)`. There are totally `r nrow(merge_unemployment)` observations and `r ncol(merge_unemployment)` variables in this dataset. The range of year is from `r min(pull(merge_unemployment, year))` to `r max(pull(merge_unemployment, year))`. Key variables should be year and month since unemployment and snps_pols_merge are joined by these two variables






