---
title: "p8105_hw2_zc2610"
author: "Zuoqiao Cui"
date: "2022-10-01"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1

```{r}
NYC_transit_data = read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line,station_name,station_latitude,station_longitude,starts_with("route"),entry, exit_only, vending, entrance_type, ada) %>% 
  mutate(
    entry = recode(entry,"YES" = "TRUE","NO" = "FALSE")
  )
```

```{r}
num_of_row = nrow(NYC_transit_data)
num_of_column = ncol(NYC_transit_data)
```


Description
1. There are these variables: `r colnames(NYC_transit_data)`
2. The process of tidying data is to import, update variable names, and selects the columns that will be used in later parts fo this problem. Then, I update `entry` from `yes` / `no` to a logical variable.
3.There are totally `r nrow(NYC_transit_data)` rows and `r ncol(NYC_transit_data)` columns.
4. I think it is not tidy enough since there are several columns regarding route and they can be combined to a column "route" with different numbers as values


Calculate the number of combinations of different station names and lines
```{r}
NYC_transit_data = NYC_transit_data %>% 
  select(station_name,line) %>% 
  distinct
  num_of_combination = nrow(NYC_transit_data)
```

Description
There are totally `r nrow(NYC_transit_data)` distinct stations




## Problem 2

Import, clean, and organize the data for Mr Trash Wheel

```{r}
MrTrashWheel_df = 
    read_excel("./data/Trash Wheel Collection Data.xlsx",sheet = "Mr. Trash Wheel",range = "A2:N550") %>% 
    janitor::clean_names() %>% 
    drop_na(dumpster) %>% 
    mutate(
      dumpster = as.character(dumpster),
      sports_balls = as.integer(round(sports_balls)),
      year = as.double(year)
    )
```
Import, clean, and organize the data for Professor Trash Wheel

```{r}
ProfessorTrashWheel_df = 
    read_excel("./data/Trash Wheel Collection Data.xlsx",sheet = "Professor Trash Wheel",range = "A2:M97") %>% 
    janitor::clean_names() %>% 
    drop_na(dumpster) %>% 
    mutate(
      dumpster = as.character(dumpster)
    )
```

Combine two sheets

```{r}
MrTrashWheel_df = MrTrashWheel_df %>% 
 mutate(type = "Mr. Trash Wheel")
ProfessorTrashWheel_df = ProfessorTrashWheel_df %>% 
  mutate(type = "Professor Trash Wheel")
TrashWheel_tidy = bind_rows(MrTrashWheel_df,ProfessorTrashWheel_df) %>% 
  janitor::clean_names() %>% 
  mutate(
    year = as.integer(year)
  )
```
Number of observations and variables in Mr Trash Wheel dataset
```{r}
num_of_MrTrashWheel_observations = nrow(MrTrashWheel_df)
num_of_MrTrashWheel_variables = ncol(MrTrashWheel_df)
num_of_MrTrashWheel_observations
num_of_MrTrashWheel_variables
```

Number of observations and variables in Mr Trash Wheel dataset
```{r}
num_of_ProfessorTrashWheel_observations = nrow(ProfessorTrashWheel_df)
num_of_ProfessorTrashWheel_variables = ncol(ProfessorTrashWheel_df)
num_of_ProfessorTrashWheel_observations
num_of_ProfessorTrashWheel_variables
```

Number of observations and variables in Mr Trash_tidy date frame
```{r}
num_of_TrashWheel_tidy_observations = nrow(TrashWheel_tidy) 
num_of_TrashWheel_tidy_variables = ncol(TrashWheel_tidy)
num_of_TrashWheel_tidy_observations
num_of_TrashWheel_tidy_variables
```
Calculate the total weight of trash collected by professor trash wheel
```{r}
professor_total_weight = sum(filter(TrashWheel_tidy,type == "Professor Trash Wheel")$weight_tons)
professor_total_weight
```

Calculate the total number of sports bals collected by Mr trash wheel
```{r}
Mr_total_sportsballs = sum((filter(TrashWheel_tidy,type == "Mr. Trash Wheel",year == 2020)$sports_balls))
Mr_total_sportsballs
```

Description
There are totally `r nrow(MrTrashWheel_df)` observations and `r ncol(MrTrashWheel_df)` variables in the Mr Trash Wheel dataset. 
There are totally `r nrow(ProfessorTrashWheel_df)` observations and `r ncol(ProfessorTrashWheel_df)` variables in the Professor Trash Wheel dataset. 
There are totally `r nrow(TrashWheel_tidy)` observations and `r ncol(TrashWheel_tidy)` variables in the final combined dataset. 
Key variables contain "dumpster","year","weight_tons","sports_balls" and so on.
The total weight of trash collected by Professor Trash Wheel is `r sum(filter(TrashWheel_tidy,type == "Professor Trash Wheel")$weight_tons)`
The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(filter(TrashWheel_tidy,type == "Mr. Trash Wheel",year == 2020)$sports_balls)`

## Problem 3

1. Separate date into year, month and day
2. Replace month number with month name
```{r}
pols_month_df = 
    read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
    janitor::clean_names() %>% 
    separate(mon,into = c("year","month","day"),sep = "-") %>% 
  mutate(
    month = recode(month, "01" = "January",
                  "02" = "February",
                  "03" = "March",
                  "04" = "April",
                  "05" = "May",
                  "06" = "June",
                  "07" = "July",
                  "08" = "Augest",
                  "09" = "September",
                  "10" = "October",
                  "11" = "November", 
                  "12" = "December"
    )
  )
pols_month_df
```
create a president column using values prez_gop and prez_dem
```{r}
pols_month_df %>% 
  mutate(
    president = ifelse(prez_gop == 1, "gop",ifelse(prez_gop == 0,"dem","special case")))
pols_month_df
```

remove prez_gop,prez_dem and day variables
```{r}
pols_month_df = pols_month_df %>% 
select(-prez_dem,-prez_gop,-day)
pols_month_df
```

separate date into year, month, day
```{r}
snps_df = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
    separate(date,into = c("month","day","year"),sep = "/")
```

arrange year,month and organize them to be leading columns
```{r}
snps_df = snps_df %>% 
mutate(
  year = as.integer(year),
  year = ifelse(year < 22,year + 2000,year + 1900)
) 
  snps_df = mutate(snps_df,
    year = as.character(year)
  ) %>% 
  arrange(year,month) %>% 
  relocate(year,month)
```
switch unemployment dataset from “wide” to “long” format
```{r}
unemployment_df = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month", 
    values_to = "percentage of unemployment") %>% 
  mutate(
    year = as.character(year)
  )

```

merge snps dataset into pols dataset and then merge unemployment dataset into them

```{r}
snps_pols_merge = left_join(pols_month_df,snps_df,by = c("year","month"))
merge_unemployment = left_join(snps_pols_merge,unemployment_df,by = c("year","month"))
```









